---
title: "Reproducible Visualization Using R"
author: "Maria Kamenetsky"
date: "January 22, 2020"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(ggplot2)
library(dplyr)

```

We start by loading several packages, including `ggplot2`, which we will use to create our reproducible visualization; and also `dplyr` ,which we will use to manipulate the data and ultimately add text to our visualization.


First, we want to read in the data. We use `read.csv()` to read in `viz_cleansamples1.csv` from our *data/* folder. We use the `str()` function to explore the structure of the dataset.

```{r}
#read in data
samples <- read.csv("../data/viz_cleansamples1.csv")
#check out data structure
str(samples)

```

## Explore the Data

Next, we want to explore the dataset. We use the `tapply()` function to calculate the summary statistics by body site.

```{r}
tapply(samples$otus, samples$bodysite, summary)
tapply(samples$otus, samples$bodysite, sum)
tapply(samples$otus, samples$sampleid, sum)
```

I want to directly add the sum total of otus by sample/body site to the plot. To do so, we first need to process the **samples** dataframe. Recall, we know the sum total of otus by sample from `tapply(samples$otus, samples$bodysite, sum)`. We will add a new variable called `sumtot` which will be the sum total by sample id. We will use the `dplyr` R package to calculate the sum total by sample id and add `sumtot` to the dataframe:

```{r}
samples <- samples %>%
    group_by(sampleid) %>%
    mutate(sumtot = sum(otus, na.rm=TRUE))
head(samples)
```



## Create Visualization

```{r}
#create base bar plot we will add to
ggplot(data=samples, aes(x=bodysite, y=otus)) +
    geom_bar(stat="identity")

#save this base plot as p1, we will add things to p1 (this will save a bit of typing)
p1 <- ggplot(data=samples, aes(x=bodysite, y=otus, fill=bodysite)) +
    geom_bar(stat="identity")
#what does p1 look like? Call the object p1
p1

#add black and white background them & change colors to goldenrod, purple, forestgreen
p1 + theme_bw() +
    scale_fill_manual(values = c("goldenrod", "purple", "forestgreen"))

#change x, y labels and add informative title to plot
p1 + theme_bw() +
    scale_fill_manual(values = c("goldenrod", "purple", "forestgreen")) +
    ylab("OTU Count") + 
    xlab("Body site") +
    ggtitle("Bar Plot of OTU Count by Body Site")

#center plot title and change legend title
p1 + theme_bw() +
    scale_fill_manual(values = c("goldenrod", "purple", "forestgreen")) +
    ylab("OTU Count") + 
    xlab("Body site") +
    ggtitle("Bar Plot of OTU Count by Body Site") +
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(fill = "Body Site")

#Change legend labels to be capitalized, separate "leftpalm" into 2 words
p1 + theme_bw() +
    scale_fill_manual(name = "Body Site",
                        labels = c("Gut", "Left Palm", "Tongue"),
                      values = c("goldenrod", "purple", "forestgreen")) +
    ylab("OTU Count") + 
    xlab("Body site") +
    ggtitle("Bar Plot of OTU Count by Body Site") +
    theme(plot.title = element_text(hjust = 0.5))

#facet by sampleid (create a multi-panel plot)
p1 + theme_bw() +
    scale_fill_manual(name = "Body Site",
                        labels = c("Gut", "Left Palm", "Tongue"),
                      values = c("goldenrod", "purple", "forestgreen")) +
    ylab("OTU Count") + 
    xlab("Body site") +
    ggtitle("Bar Plot of OTU Count by Body Site") +
    theme(plot.title = element_text(hjust = 0.5)) +
    facet_wrap(.~sampleid)


#add text to each panel
p1 + theme_bw() + 
    scale_fill_manual(name = "Body Site",
                        labels = c("Gut", "Left Palm", "Tongue"),
                      values = c("goldenrod", "purple", "forestgreen")) +
    ylab("OTU Count") + 
    xlab("Body site") +
    ggtitle("Bar Plot of OTU Count by Body Site") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_text(aes(label=sumtot), y=6000) +
    facet_wrap(.~sampleid) 

#save my visualization
ggsave("myviz.png", path = "../figures/")
```




























